### 2021-06-10 papers/thesis/full/ch4-spherical-caps/spherical-caps.tex
Added index of notation. Uses makeindex.

By construction, $\{\scopnkia\}$ are orthogonal with respect to the inner product
\bseqn{
	\ip< p, \: q >_{\scopa} &:= \int_\Omega p(\xvec, z) \: q(\xvec, z) \: \genjacw^{(a,0)}(z) \: \D A \\
	&= \int_\alpha^1 \int_0^{2\pi} p(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: q(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: \D \theta \: \genjacw^{(a,0)}(z) \: \D z,
}
with
\bseqnnumber{
	\norm{\scopnkia}_{\scopa}^2 &:= \ip< \scopnkia, \: \scopnkia >_{\scopa} = \pi \: \normgenjac^{(a, 2k)}. \label{eqn:sc:normscop}
}
A method for obtaining explicit recurrence coefficients and evaluating integrals for the weight function $\genjacw^{(a,b)}(x)$ was established in \bsrefchapter{CHAPTER:diskslice} (see also \cite{snowball2019sparse}). The weight is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}.

@@

The recurrences in \bsreflemma{lemma:sc:Qrecurrence} lead to (block) Jacobi matrices that correspond to multiplication by $x$, $y$ and $z$. In later sections we will use an ordering of the OPs so that they are grouped by Fourier mode $k$, which is convenient for the application of differential and other operators to the vector of coefficients of a given function's expansion (some operators will exploit this ordering for operators where Fourier modes do not interact, and thus will be block-diagonal). Before that, the ordering we will use in the remainder of this section is convenient for establishing Jacobi operators for multiplication by $x$, $y$ and $z$, and hence building the OPs and importantly obtaining the associated \textit{recurrence coefficient matrices} necessary for efficient function evaluation using the Clenshaw algorithm. In practice, it is simply a matter of converting coefficients between the two orderings. To this end, we define our OP-building ordering as follows. For $n=0,1,2,\dots$:
\bseqn{
	\bigscoptna := 
		\begin{pmatrix}
			\scopa_{n,0,0}(x,y,z) \\
			\scopa_{n,1,0}(x,y,z) \\
			\scopa_{n,1,1}(x,y,z) \\
			\vdots \\
			\scopa_{n,n,0}(x,y,z) \\
			\scopa_{n,n,1}(x,y,z)
		\end{pmatrix} \in \R^{2n+1}, 
	\quad \quad 
	\bigscopta := 
		\begin{pmatrix}
			\bigscopta_0 \\
			\bigscopta_1 \\
			\bigscopta_2 \\
			\vdots \\
		\end{pmatrix}
}
and set $ J_x^{(a)},  J_y^{(a)},  J_z^{(a)}$ as the Jacobi matrices corresponding to
\bseqnnumber{
	J_x^{(a)} \: \bigscopta(x,y,z) = x \: \bigscopta(x,y,z), \nonumber \\
	J_y^{(a)} \: \bigscopta(x,y,z) = y \: \bigscopta(x,y,z), \label{eqn:sc:jacobimatricesdefinitionalt} \\
	J_z^{(a)} \: \bigscopta(x,y,z) = z \: \bigscopta(x,y,z). \nonumber
}
where
\bseqn{
	J_{x/y/z}^{(a)} &= 
		\begin{pmatrix}
			B^{(a)}_{x/y/z, 0} & A^{(a)}_{x/y/z, 0} & & & & \\
			C^{(a)}_{x/y/z, 1} & B^{(a)}_{x/y/z, 1} & A^{(a)}_{x/y/z, 1} & & & \\
			& C^{(a)}_{x/y/z, 2} & B^{(a)}_{x/y/z, 2} & A^{(a)}_{x/y/z, 2} & & & \\
			& & C^{(a)}_{x/y/z, 3} & \ddots & \ddots & \\
			& & & \ddots & \ddots & \ddots \\
		\end{pmatrix}.
}
While $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are not Jacobi matrices in the classical sense, they are block Jacobi matrices. However, it is useful to simply label them as such. They are once again in fact banded-block-banded matrices -- see \bsrefdef{def:bandedblockbanded}. More specifically, each Jacobi matrix is block-tridiagonal (block-bandwidths $(1,1)$).

@@

In this section we will derive the entries of spherical partial differential operators applied to our basis, demonstrating their sparsity in the process. To this end, as alluded to in \bsrefsection{section:sc:jacobimats}, we introduce new notation for a different ordering of the OP vector, in order to exploit the orthogonality the polynomials $\chki$ will bring and thus ensure the operators will be block-diagonal. Let $N \in \N$ and define:
\bseqnnumber{
	\bigscopNka &:= 
		\begin{pmatrix}
			\scopa_{k,k,0}(x,y,z) \\
			\scopa_{k,k,1}(x,y,z) \\
			\vdots \\
			\scopa_{N,k,0}(x,y,z) \\
			\scopa_{N,k,1}(x,y,z)
		\end{pmatrix} \in \R^{2(N-k+1)},  \quad k = 1,\dots,N, \label{eqn:sc:OPdefNka} \\ 
	\bigscopa_{N,0} &:= 
		\begin{pmatrix}
			\scopa_{0,0,0}(x,y,z) \\
			\vdots \\
			\scopa_{N,0,0}(x,y,z) \\
		\end{pmatrix} \in \R^{N+1}, \label{eqn:sc:OPdefN0a} \\
	\bigscopNa &:= 
		\begin{pmatrix}
			\bigscopa_{N,0} \\
			\vdots \\
			\bigscopa_{N,N}
		\end{pmatrix} \in \R^{(N+1)^2} \label{eqn:sc:OPdefNa}
}

We further denote the weighted set of OPs on $\Omega$ by 
\bseqn{
	\bigWNa(x,y,z) := \genjacw^{(a,0)}(z) \: \bigscopNa(x,y,z),
}

>>>>

By construction, $\{\scopnkia\}$ are orthogonal with respect to the inner product \nomenclature[innerQ]{$\ip< p, \: q >_{\scopa}$}{A multidimensional inner product over $\Omega$ with the weight $\Wa$ associated with the OPs $\{\scopnkia\}$}
\bseqn{
	\ip< p, \: q >_{\scopa} &:= \int_\Omega p(\xvec, z) \: q(\xvec, z) \: \genjacw^{(a,0)}(z) \: \D A \\
	&= \int_\alpha^1 \int_0^{2\pi} p(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: q(\rho(z) \cos\theta, \rho(z) \sin\theta, z) \: \D \theta \: \genjacw^{(a,0)}(z) \: \D z,
}
with \nomenclature[normQ]{$\norm{f}_{\scopa}$}{The norm of a function f, defined by the square root of the inner product over $\Omega$ given by the weight $\Wa$ associated with the OPs $\{\scopnkia\}$}
\bseqnnumber{
	\norm{\scopnkia}_{\scopa}^2 &:= \ip< \scopnkia, \: \scopnkia >_{\scopa} = \pi \: \normgenjac^{(a, 2k)}. \label{eqn:sc:normscop}
}
A method for obtaining explicit recurrence coefficients and evaluating integrals for the weight function $\genjacw^{(a,b)}(x)$ was established in \bsrefchapter{CHAPTER:diskslice} (see also \cite{snowball2019sparse}). The weight is in fact semi-classical, and is equivalent to a generalized Jacobi weight \cite[\S5]{magnus1995painleve}.

@@

The recurrences in \bsreflemma{lemma:sc:Qrecurrence} lead to (block) Jacobi matrices that correspond to multiplication by $x$, $y$ and $z$. In later sections we will use an ordering of the OPs so that they are grouped by Fourier mode $k$, which is convenient for the application of differential and other operators to the vector of coefficients of a given function's expansion (some operators will exploit this ordering for operators where Fourier modes do not interact, and thus will be block-diagonal). Before that, the ordering we will use in the remainder of this section is convenient for establishing Jacobi operators for multiplication by $x$, $y$ and $z$, and hence building the OPs and importantly obtaining the associated \textit{recurrence coefficient matrices} necessary for efficient function evaluation using the Clenshaw algorithm. In practice, it is simply a matter of converting coefficients between the two orderings. To this end, we define our OP-building ordering as follows. For $n=0,1,2,\dots$: \nomenclature[Qtildevectora]{$\bigscopta$}{Vector of the spherical cap OPs ordered by degree, with parameter $a$} \nomenclature[Qtildevectorna]{$\bigscoptna$}{(Sub-)Vector of the spherical cap OPs of degree $n$, with parameter $a$}
\bseqn{
	\bigscoptna := 
		\begin{pmatrix}
			\scopa_{n,0,0}(x,y,z) \\
			\scopa_{n,1,0}(x,y,z) \\
			\scopa_{n,1,1}(x,y,z) \\
			\vdots \\
			\scopa_{n,n,0}(x,y,z) \\
			\scopa_{n,n,1}(x,y,z)
		\end{pmatrix} \in \R^{2n+1}, 
	\quad \quad 
	\bigscopta := 
		\begin{pmatrix}
			\bigscopta_0 \\
			\bigscopta_1 \\
			\bigscopta_2 \\
			\vdots \\
		\end{pmatrix}
}
and set $ J_x^{(a)},  J_y^{(a)},  J_z^{(a)}$ as the Jacobi matrices corresponding to
\bseqnnumber{
	J_x^{(a)} \: \bigscopta(x,y,z) = x \: \bigscopta(x,y,z), \nonumber \\
	J_y^{(a)} \: \bigscopta(x,y,z) = y \: \bigscopta(x,y,z), \label{eqn:sc:jacobimatricesdefinitionalt} \\
	J_z^{(a)} \: \bigscopta(x,y,z) = z \: \bigscopta(x,y,z). \nonumber
}
where
\bseqn{
	J_{x/y/z}^{(a)} &= 
		\begin{pmatrix}
			B^{(a)}_{x/y/z, 0} & A^{(a)}_{x/y/z, 0} & & & & \\
			C^{(a)}_{x/y/z, 1} & B^{(a)}_{x/y/z, 1} & A^{(a)}_{x/y/z, 1} & & & \\
			& C^{(a)}_{x/y/z, 2} & B^{(a)}_{x/y/z, 2} & A^{(a)}_{x/y/z, 2} & & & \\
			& & C^{(a)}_{x/y/z, 3} & \ddots & \ddots & \\
			& & & \ddots & \ddots & \ddots \\
		\end{pmatrix}.
}
While $J_x^{(a)}, J_y^{(a)}, J_z^{(a)}$ are not Jacobi matrices in the classical sense, they are block Jacobi matrices. However, it is useful to simply label them as such. They are once again in fact banded-block-banded matrices -- see \bsrefdef{def:bandedblockbanded}. More specifically, each Jacobi matrix is block-tridiagonal (block-bandwidths $(1,1)$).

@@

In this section we will derive the entries of spherical partial differential operators applied to our basis, demonstrating their sparsity in the process. To this end, as alluded to in \bsrefsection{section:sc:jacobimats}, we introduce new notation for a different ordering of the OP vector, in order to exploit the orthogonality the polynomials $\chki$ will bring and thus ensure the operators will be block-diagonal. Let $N \in \N$ and define: \nomenclature[QvectorNa]{$\bigscopNa$}{Vector of the spherical cap OPs up to degree $N$, grouped by order $k$, with parameter $a$} \nomenclature[QvectorNka]{$\bigscopNka$}{(Sub-)Vector of the spherical cap OPs of order $k$ up to degree $N$, with parameter $a$}
\bseqnnumber{
	\bigscopNka &:= 
		\begin{pmatrix}
			\scopa_{k,k,0}(x,y,z) \\
			\scopa_{k,k,1}(x,y,z) \\
			\vdots \\
			\scopa_{N,k,0}(x,y,z) \\
			\scopa_{N,k,1}(x,y,z)
		\end{pmatrix} \in \R^{2(N-k+1)},  \quad k = 1,\dots,N, \label{eqn:sc:OPdefNka} \\ 
	\bigscopa_{N,0} &:= 
		\begin{pmatrix}
			\scopa_{0,0,0}(x,y,z) \\
			\vdots \\
			\scopa_{N,0,0}(x,y,z) \\
		\end{pmatrix} \in \R^{N+1}, \label{eqn:sc:OPdefN0a} \\
	\bigscopNa &:= 
		\begin{pmatrix}
			\bigscopa_{N,0} \\
			\vdots \\
			\bigscopa_{N,N}
		\end{pmatrix} \in \R^{(N+1)^2} \label{eqn:sc:OPdefNa}
}

We further denote the weighted set of OPs on $\Omega$ by \nomenclature[WvectorNa]{$\bigWNa$}{Vector of the weighted spherical cap OPs up to degree $N$, grouped by order $k$, with parameter $a$}
\bseqn{
	\bigWNa(x,y,z) := \genjacw^{(a,0)}(z) \: \bigscopNa(x,y,z),
}
